An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-05-10 EconGrowth.Rnw
Added details to energy analysis. Added a paragraph on organization of the paper. Added references.

@@

\noindent where $C_{fuel}$ is the consumption rate of the fuel (in units of short tons/year for coal, barrels/year for oil and petroleum, ft$^3$/year for dry natural gas and natural gas plant liquids, and m$^3$/year for biomass/fuel wood) and $h_{{fuel}}$ is the heat content of the fuel (in kJ/short ton for coal, kJ/barrel for oil, kJ/ft$^3$ for dry natural gas and natural gas plant liquids, and kJ/kg for biomass/fuel wood). Fuel wood density (166.67~kg/m$^3$ **** Caleb: verify this number and source ****) was obtained from the \citet{FAO:2013a}.

Yearly consumption rates ($C_{fuel}$) of coal, oil and petroleum, dry natural gas and natural gas plant liquids, and biofuels (ethanol and biodiesel) were obtained from the \citet{EIA:2013a} for all economies. Yearly average heat content data for coal, oil, and natural gas and natural gas plant liquids was obtained from the EIA for all economies. We used 26.8~MJ/kg \citep[Table~A-27]{Cengel:2011ww} and 37.8~MJ/kg \citep[Table~7.3]{Gupta:2010vg} for the heat content of ethanol and corn-based biodiesel, respectively. Yearly consumption and heat content data for biomass/fuel wood was obtained from the ``Fuelwood'' time series of the FAO Statistics Database \citep{FAO:2013a}.

@@

Exergy ($X$), expressed in TJ/year, was calculated from the thermal energy ($Q$) values discussed in Section \ref{sec:Q_Data}. For fuels, exergy was calculated by a unitless exergy-to-energy allocation factor ($f_{exergy-allocation}$) from \citet[Table~A-2]{Ayres:2003ec} and shown in Table \ref{tab:Exergy_Allocation}.

@@

\begin{table} \caption{Exergy allocation factors.} \label{tab:Exergy_Allocation} 
  \begin{center}
    \begin{tabular}{r l} 
      Fuel                                    & $f_{exergy-allocation}$ \\
      \hline
      Coal                                    & 1.088    \\
      Oil and petroleum                       & 1.073   \\      
      Dry natural gas and                     & \multirow{2}{*}{1.04}    \\
      natural gas liquids                     &         \\
      Biomass/fuel wood                       & 1.15   \\
      Nuclear                                 & 1.0   \\
      Renewables                              & 1.0   \\
    \end{tabular}
  \end{center}
\end{table}

\noindent $f_{exergy-allocation}$ is the ratio of chemical exergy to the higher heating value for each fuel. $f_{exergy-allocation}$ for each fuel was assumed to be constant through time across all economies. For renewable and nuclear energy sources, the electricity generated is already pure exergy, so $f_{exergy-allocation}$ is unity.

>>>>

The remainder of this paper is organized as follows. Section \ref{sec:Coordinates} discusses the coordinates of analysis described above. Section \ref{sec:Sources_of_Data} details sources of data and calculations to generate time series for economic output and factors of production. The procedures for and results from parameter estimation for the various production functions are discussed in Section \ref{sec:Parameter_Estimation}. Section \ref{sec:Implications} discusses implications of the parameter estimation results, and Section \ref{sec:Conclusion} concludes. An appendix provides details of all statistical results.

@@

\noindent where $C_{fuel}$ is the consumption rate of the fuel (in units of short tons/year for coal, barrels/year for oil and petroleum, ft$^3$/year for dry natural gas and natural gas plant liquids, and m$^3$/year for biomass/fuel wood) and $h_{{fuel}}$ is the heat content of the fuel (in kJ/short ton for coal, kJ/barrel for oil, kJ/ft$^3$ for dry natural gas and natural gas plant liquids, and kJ/kg for biomass/fuel wood). Fuel wood density (166.67~kg/m$^3$ **** Caleb: verify this number and source ****) was calculated from ***** and *****, both obtained from the \citet{FAO:2013c}.

Yearly consumption rates ($C_{fuel}$) of coal, oil and petroleum, dry natural gas and natural gas plant liquids, and biofuels (ethanol and biodiesel) were obtained from the \citet{EIA:2013a} for all economies. Yearly average heat content data for coal, oil, and natural gas and natural gas plant liquids was obtained from the EIA for all economies. We used lower (net) heating values of 26.8~MJ/kg \citep[Table~A-27]{Cengel:2011ww} and 37.8~MJ/kg \citep[Table~7.3]{Gupta:2010vg} for the heat content of ethanol and corn-based biodiesel, respectively. Yearly consumption and heat content data for biomass/fuel wood was obtained from the ``Fuelwood'' time series of the FAO Statistics Database \citep{FAO:2013a}.

@@

Exergy ($X$), expressed in TJ/year, was calculated by multiplying the thermal energy ($Q$) values discussed in Section \ref{sec:Q_Data} by a unitless exergy-to-thermal-energy allocation factor ($f_{exergy-allocation}$) shown in Table \ref{tab:Exergy_Allocation}.

@@

\begin{table} \caption{Exergy allocation factors.} \label{tab:Exergy_Allocation} 
  \begin{center}
    \begin{tabular}{r l} 
      Fuel                                    & $f_{exergy-allocation}$ \\
      \hline
      Coal                                    & 1.088    \\
      Oil and petroleum                       & 1.073   \\      
      Dry natural gas and                     & \multirow{2}{*}{1.04}    \\
      natural gas liquids                     &         \\
      Corn-based biodiesel                    & 1.07  \\
      Biomass/fuel wood                       & 1.15   \\
      Ethanol biofuel                         & 1.099 \\
      Nuclear, wind,                          & \multirow{2}{*}{1.0}   \\
      solar PV, and hydro                     &                        \\
    \end{tabular}
  \end{center}
\end{table}

\noindent For carbon-based energy sources, $f_{exergy-allocation}$ is the ratio of chemical exergy to lower (net) heating value. The exergy allocation factors for coal, oil, natural gas, biodiesel, and biomass/fuel wood were obtained from \citet[Table~A-2]{Ayres:2003ec}. The exergy allocation ratio for liquid ethanol was calculated from the standard molar chemical exergy of 1357.7~kJ/mol \citep[Table~II]{Szargut:1988wv} and its lower (net) heating value of 26.8~MJ/kg \citep[Table~A-27]{Cengel:2011ww}. $f_{exergy-allocation}$ for each fuel was assumed to be constant through time across all economies. For nuclear, wind, solar photovoltaic, and hydro energy sources, the electricity generated is already pure exergy, so $f_{exergy-allocation}$ is unity.
